Introduction to Lattice-based Attacks\\``Cryptography''

### 2021-06-04 paper.tex
Add references for ECDSA attacks

\title{Introduction to Lattices Attacks\\``Cryptography''}

@@

Kelby Ludwig explains an intuition behind every steps of the algorithm \cite{lllintuition}, and Oded Regev
gives a more rigorous explanation of the algorithm and the properties \cite{lllexplanation}.

\section{Applications}

There are many applications of LLL

@@

\begin{itemize}
    \item If $m < N^{\frac{1}{e}}$, then $m^e < N$ and the modulo operation is not applied and we only need to find the $e$th root of $c$ over the
        integers to find $m$.
    \item If $q = p + x$ where $x$ is small enough than it's easy to recover the factors of $N$ computing $A = \sqrt{N}$ and compute
            $p = A - x$ for different value of $x$ until $N \mod p = 0$, then we have found $q = N / p$.
    \item If $d$ is too small (for example to accelerate decryption) then Wiener found a way to recover the private key $d$ \cite{wiener}.
    \item Other attacks can be found on the excellent survey by Dan Boneh \cite{boneh99twentyyears}.
\end{itemize}

@@

We start introducing the main ideas of Coppersmith's attack \cite{coppersmith96} and the math behind it.
The attack can be used against \textbf{RSA} when a small $e$ is used, but also
to factorize the modulus. In 2017 the attack was used to exploit
\href{https://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2017-15361}{CVE-2017-15361} also named \textbf{ROCA}
(Return Of Coppersmith's Attack) \cite{roca}.

@@

\subsection{Real World case study: ROCA}

\textbf{ROCA}\cite{roca} is a vulnerability discovered in the 2017 in a software library \textit{RSALib} in the generation
of the public modulus for \textbf{RSA}. The attack is believed to affects millions of smart cards. The vulnerbality
resides in the generation of prime numbers, since embedded devices don't have lots of computational resources a fast way
was needed. Every prime generated has the following structure:

@@

Where $k, a$ are random values different for every prime (unkown), $g = 65537$ and $M$ is known and it's a \textit{primorial} number.
A \textit{primorial} number $M$ is the product of the the first $n$ successive primes

@@

It's possible to compute the discrete logarithm of $g^c \mod M$ since the size of $|M|$ is smooth with
Pohlig-Hellman algorithm \cite{pohlig}. With $c$ we can set a bound to the value of $a$ since

@@

\begin{algorithm}[H]
    \vspace*{5px}
    \textbf{Input: } $N, M, g, ord, X$ \; \\
    $c = \log_{g}N \mod M$ \;\\
    \For{$a \gets (\frac{c}{2}, \frac{c + ord}{2})$}{
        \vspace*{5px}
        $f(x) = xM + (g^a \mod M)$ \; \\
        $g(x) = \frac{1}{M}f(x)$ [make monic, the roots are the same] \; \\
        $k = $ coppersmith($g(x), N, X$) [find k] \; \\ 
        $p = kM + (g^a \mod M)$  [candidate] \; \\
        \If{$N \mod p = 0$}{
            return $p$ \;
        }
    }
\end{algorithm}

In practice however the bound of $a$ is too big, so the authors found a way the problem to a smaller $M'$ since the
bitsize of $M$ are more than sufficient for making the attack works with Coppersmith's algorithm. Their optimization
find a new $M'$ such that

@@

We were able to reproduce the attack in sagemath against \textbf{RSA}-512, and we used as $M'$ the value
that Bruno Produit \cite{rocaopt} made available in his thesis and David Wong's implementation of Coppersmith's attack \cite{wong2015}.

@@

\textbf{ECDSA} is a variant of the Digital Signature Algorithm (\textbf{DSA}) which uses elliptic curve cryptography.
To digitally sign a message we have 3 public parameters

@@

We also need to create a private key $d \in [1, n-1]$ and public key $Q = dG$. To digitally \textbf{sign} a message $m$:

\begin{enumerate}
    \item Compute $h = $ HASH$(m)$ where HASH is a cryptographic hash functions.
    \item Select a random integer $k \in [1, n-1]$.
    \item Calculate $P = kG = (x_1, y_1)$ and set $r = (x_1)$.
    \item Compute $s = k^{-1}(h + dr) \mod n$, if $s = 0$ repeat the steps.
    \item The signature is composed by $(r, s)$.
\end{enumerate}

@@

\begin{itemize}
    \item If the same $k$ is used to generate two different signatures it's possible to recover the secret key $d$. This was a real bug 
        discovered in the Playstation 3.
    \item It's possible to recover $d$ if the parameter $k$ is not generated with a cryptographically secure pseudo random generator.
    \item If the elliptic curve used is not standardized (custom) then it's possible that the discrete logarithm is easily solvable.
\end{itemize}

@@

@@

Let $p = \texttt{0xffffffffffffd21f}$ and let $E: y^2 = x^3 + 3$ be an elliptic curve over $\F_p$ with the generator
$G = (\texttt{0xcc3b3d1a0c4938ef}, \texttt{0x4ab35ff66f8194fa})$ of order $n = \texttt{0xfffffffefa23f437}$.

@@

Same as before this lattice contains $v = (-k_1, -k_2, \ldots, k_m, K)$ and with probability TODO we can find that vector after applying LLL.

@@

I was able to recover the nonces with the $10$ MSB bits known on the curve \texttt{secp256r1} given $100$ signatures, but it's also possible
to recover the nonces with less known bits as shown in \href{https://eprint.iacr.org/2020/1540}{paper}.

>>>>

\title{Introduction to Lattice-based Attacks\\``Cryptography''}

@@

Kelby Ludwig explains an intuition behind every steps of the algorithm \cite{lllintuition}, and Oded Regev
gives a more rigorous explanation of the algorithm and the properties \cite{lllexplanation}.\\

Some of the applications of the algorithm are the following:

@@

\begin{itemize}
    \item If $m < N^{\frac{1}{e}}$, then $m^e < N$ and the modulo operation is not applied and we only need to find the $e$th root of $c$ over the
        integers to find $m$.
    \item If $q = p + x$ where $x$ is small enough than it's easy to recover the factors of $N$ computing $A = \sqrt{N}$ and compute
            $p = A - x$ for different value of $x$ until $N \mod p = 0$, then we have found $q = N / p$.
    \item It's possible to recover $d$ if it is too small (for example to accelerate decryption) using the Wiener's attack \cite{wiener}.
    \item Other attacks can be found on the excellent survey by Dan Boneh \cite{boneh99twentyyears}.
\end{itemize}

@@

We start introducing the main ideas of Coppersmith's attack \cite{coppersmith96} and the math behind it.
The attack can be used against \textbf{RSA} when a small $e$ is used, but also
to factorize the modulus if some bits of $p$ or $q$ are known. In 2017 the attack was used to exploit
\href{https://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2017-15361}{CVE-2017-15361} also named \textbf{ROCA}
(Return Of Coppersmith's Attack) \cite{roca}.

@@

\subsection{Real world case study: ROCA}

\textbf{ROCA}\cite{roca} is a vulnerability discovered in 2017 in a software library \textit{RSALib} in the generation
of the public modulus for \textbf{RSA}. The attack is believed to affects millions of smart cards. The vulnerbality
resides in the generation of prime numbers, since embedded devices don't have lots of computational resources a fast way
was needed. Every prime generated has the following structure:

@@

Where $k, a$ are random values different for every prime (unkown), $g = 65537$ and $M$ it's a \textit{primorial} number.
A \textit{primorial} number $M$ is the product of the the first $n$ successive primes

@@

It's possible to compute the discrete logarithm of $g^c \mod M$ since the size of $|M|$ is smooth with
Pohlig-Hellman algorithm \cite{pohlig}. With $c$ we can set the bound to the value of $a$ since

@@

\begin{algorithm}[H]
    \vspace*{5px}
    \textbf{Input: } $N, M, g, ord, X$ \; \\
    $c = \log_{g}N \mod M$ \;\\
    \For{$a \gets (\frac{c}{2}, \frac{c + ord}{2})$}{
        \vspace*{5px}
        $f(x) = xM + (g^a \mod M)$ \; \\
        $g(x) = \frac{1}{M}f(x)$ [make monic, the roots are the same] \; \\
        $k = $ coppersmith($g(x), N, X$) [find k] \; \\ 
        $p = kM + (g^a \mod M)$  [candidate] \; \\
        \If{$N \mod p = 0$}{
            return $p$ \;
        }
    }
    \caption{Exploit idea}
\end{algorithm}

In practice however the bound of $a$ is too big, so the authors found a way to move the same problem to a smaller $M'$ since the
bitsize of $M$ are more than sufficient for making the attack works with Coppersmith's algorithm. Their optimization
find a new $M'$ such that

@@

We were able to reproduce the attack in sagemath against \textbf{RSA}-$512$, and we used as $M'$ the value
that Bruno Produit \cite{rocaopt} made available in his thesis and David Wong's implementation of Coppersmith's attack \cite{wong2015}.

@@

\textbf{ECDSA} is a variant of the Digital Signature Algorithm (\textbf{DSA}) which uses elliptic curve cryptography.
There are 3 public parameters:

@@

We also need to create a private key $d \in [1, n-1]$ and public key $Q = dG$. To \textbf{sign} a message $m$:

\begin{enumerate}
    \item Compute $h = $ MSB(HASH$(m))$ where HASH is a cryptographic hash function, for example SHA-256, and MSB take the $n$ most significant bits.
    \item Select a random integer $k \in [1, n-1]$ (nonce).
    \item Calculate $P = kG = (x_1, y_1)$ and set $r = (x_1)$ if $x_1 \neq 0$, otherwise repeat the steps.
    \item Compute $s = k^{-1}(h + dr) \mod n$, if $s = 0$ repeat the steps.
    \item The signature is composed by $(r, s)$.
\end{enumerate}

@@

\begin{itemize}
    \item If the same $k$ is used to generate two different signatures it's possible to recover the secret key $d$. This was a real bug 
        discovered in the Playstation 3.
    \item It's possible to recover $d$ if the parameter $k$ is not generated with a cryptographically secure pseudo random generator.
    \item If the elliptic curve used is not standardized (custom) then it's possible that the discrete logarithm is easily solvable
        \cite{giuliani99attackson}, \cite{mov} \cite{smartass}.
\end{itemize}

@@

The lattice-based attacks presented here are based on the \textit{Hidden Number Problem} introduced
by Boneh and Venkatesan \cite{hnp} that was in a second moment generalized to prove that the most significan bits
of $k$ can potentially leak the entire nonce \cite{ecdsa_nonce}.

@@

\textbf{Problem setup}.
Let $p = \texttt{0xffffffffffffd21f}$ and let $E: y^2 = x^3 + 3$ be an elliptic curve over $\F_p$ with the generator
$G = (\texttt{0xcc3b3d1a0c4938ef}, \texttt{0x4ab35ff66f8194fa})$ of order $n = \texttt{0xfffffffefa23f437}$.

@@

Same as before this lattice contains $v = (-k_1, -k_2, \ldots, k_m, K)$ that can be found using LLL.

@@

Without any optimization we were able to recover the nonces with the $10$ MSB bits known on the curve \texttt{secp256r1} given $100$ signatures,
but it's also possible to recover the nonces with less known bits as shown in the results obtained by Albrecht and Heninger \cite{ecdsa_result}.
